An Open Repository of Real-Time COVID-19 Indicators

### 2021-06-16 indicators/paper/paper.tex
Cite recent preprint on backfill in COVID
Add a new perspective

Research on data revisions in the context of influenza-like illness has shown
that backfill can significantly alter forecast performance \cite{Brooks:2018,
  Reich:2019}, and that careful training on preliminary data can reduce this
influence \cite{Brooks:2020}. We also investigate this in our companion paper on
forecasting, where we observe that training and validating models on finalized 
data yields overly optimistic estimates of true test-time performance.

% TODO: Cite this? https://arxiv.org/abs/2106.04420

@@

Auxiliary signals (outside of the standard public health reporting streams) can
serve as indicators of COVID activity, but they can also illuminate other
aspects of the pandemic. To understand the effect of the pandemic on society and
predict its future spread requires a comprehensive view of not just cases and
deaths, but public behavior and beliefs. Measures of mobility reflect social
distancing policies and the potential rate of COVID spread; medical claims data
reflects healthcare-seeking behavior and how it may be affected by local case
rates; measures of COVID vaccine acceptance can guide outreach efforts
\cite{TODO}.

% TODO

% AR: Potentially interesting things to show or comment on:
% - Medical claims data can give info about medical-seeking behavior and how it
% varies with caseload -- e.g. are people deterred from seeking medical care
% because of the high case load?
% - Mobility changes
% - Map of one of the unique signals during an outbreak, e.g. Michigan in April
% 2021

% RJT: @AR Since we already stuff on claims, can we do: mobility, masks, vaccine
% uptake or hesitancy, distancing behaviors, how worried are you?
% How about this (in each case, each point in said plot is a county's signal
% values averaged over some period)
% - Bivariate plot of mask use versus how worried are you
% - Bivariate plot of of how worried are you vs case rates
%- Bivariate plot of of how worried are you vs hospitalizations
% I would be curious to see all three.  And perhaps color code the points by
% region to highlight regional trends

>>>>

Research on data revisions in the context of influenza-like illness has shown
that backfill can significantly alter forecast performance \cite{Brooks:2018,
  Reich:2019}, and that careful training on preliminary data can reduce this
influence \cite{Brooks:2020}. Recent research has shown similar results for
COVID-19 forecasts \cite{Kamarthi:2021}. We also investigate this in our
companion paper on forecasting, where we observe that training and validating
models on finalized data yields overly optimistic estimates of true test-time
performance.

@@

Auxiliary signals (outside of the standard public health reporting streams) can
serve as indicators of COVID activity, but they can also illuminate other
aspects of the pandemic. To understand the effect of the pandemic on society and
predict its future spread requires a comprehensive view of not just cases and
deaths, but public behavior and beliefs. Measures of mobility reflect social
distancing policies and the potential rate of COVID spread; medical claims data
reflects healthcare-seeking behavior and how it may be affected by local case
rates; measures of COVID vaccine acceptance can guide outreach efforts.

\begin{figure}
  \centering
  \includegraphics[width=\columnwidth]{fig/mobility-drop.pdf}
  \caption{Percentage point increase in SafeGraph estimates of the percentage of
  people staying completely at home, based on aggregated mobile device data,
  between March 1, 2020 and April 15, 2020.}
  \label{fig:mobility-drop}
\end{figure}

As an illustration, Figure~\ref{fig:mobility-drop} maps the sharp increase in
rates of people staying at home between March 1 and April 15, 2020. Using
SafeGraph data on the fraction of mobile devices included in SafeGraph's panel
that did not leave the immediate area of their home, it illustrates the sharp
drop in travel and work outside the home that occurred in the early stages of
the pandemic. It also shows that this drop was much more pronounced in some
states than others, enabling analysis of policy impacts and disease spread.
Similar maps can be quickly constructed for any signal in the COVIDcast API.
